### 2022-09-19 manuscript.Rmd
tabellerna med nu.
det gör lite ont i ögonen att titta på dem just nu tyvärr, formatering har varit prio 2 jag ville bara presentera datan från förra veckan för diskussionen idag 19/9

---
title: "Quantifying unmet ICU beds in tertiary hospitals in urban India using Prospective Score Matching "
output: html_document
author: "Maria Eriksson"
bibliography: bibliography.bib
csl: vancouver.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
table_nums <- captioner::captioner(prefix = "Tab.")
```

<!-- This is an R Markdown document. Markdown is a simple formatting
syntax for authoring HTML, PDF, and MS Word documents. For more
details on using R Markdown see http://rmarkdown.rstudio.com -->

<!-- This is a comment and will not be present in the compiled
document. Try "knitting" this document by clicking the knit button up
to the left, or by running `rmarkdown::render("manuscript.Rmd")` in
the console and you should only see headings -->

<!-- Structure and most comments are from the Strengthening the
reporting of observational studies in epidemiology (STROBE) statement,
see https://www.strobe-statement.org/. For more explanations see the
paper Strengthening the Reporting of Observational Studies in
Epidemiology (STROBE):explanation and elaboration.  -->

<!-- Indicate the study’s design with a commonly used term in the
title, for example: "Associations between resuscitation interventions
and opportunities for improvement in adult trauma patients: A cohort
study" -->

Abstract
========

<!-- Provide in the abstract an informative and balanced summary of
what was done and what was found. Not more than 300 words. -->

Background
----------

Methods
-------

Results
-------

Conclusion
----------

Introduction
============

<!-- Explain the scientific background and rationale for the
investigation being reported. End by stating the aim or specific
objectives, including any prespecified hypotheses. The introduction
should not be longer than 1-1.5 pages, typically 4-5 paragraphs, each
6-8 lines long. Support every statement with a citation -->
Trauma/Injury is a substantial global health problem that disproportionately affects low and middle-income countries (LMIC), resulting in significant mortality and morbidity@traumaLMIC. The annual mortality from injury in India amount to one million casualties @onemilliondead. Half of trauma mortality in India occur in-hospital, and in-hospital trauma mortality in Indian hospitals was double that of high-income countries @indiantrauma. Health care professionals in many LMIC testify that their medical assessment oftentimes is that a trauma patient should be admitted to the ICU, but aren’t because such resources are scarce, overloaded and unavailable. Their medical judgement is that the lives of these patients were salvageable, and that these patients died as a direct result of not being admitted to the ICU@stuffLMICdoctorssay. Globally, an estimated two million lives could be saved annually, if the appropriate and timely trauma care provided by hospitals in many high-income countries were available and provided in low to-middle-income countries (LMICs) @livessaved

In order to improve intra-hospital care, trauma quality improvement programs (TQIP) have been developed. The aim of TQIPs is to ensure equivalent care and improve trauma care by developing best practice guidelines@tqip. TQIPs utilizes a retrospective cohort study, analysing the current of trauma care patients and post-fact attempting to determine the appropriate level of care for each patient. From these analyses many aspects of trauma care can be improved and optimized, such as determining the appropriate amount of ICU beds and other trauma care resources @tqip.

Existing TQIP often originate in high-income countries, and are either not applicable nor easily transferred to the hospital conditions of low and middle income countries (LMIC). This study aims to make an TQIP assessment in an Indian context to fill part of that knowledge gap, and to test whether there is merit in the testimony of the Indian colleagues. The study shall propose a scope of the unmet need for ICU beds for trauma patients at tertiary hospitals in urban India. If this analysis indeed indicate a shortage, it can be used to motivate extending ICU resources and reduce preventable trauma morbidity and mortality.

Trauma disproportionately affects young, otherwise healthy individuals. They are often the primary breadwinners for their families and therefore the detrimental impact of trauma go beyond mortality or substantial loss of disease-free years of life for the affected individuals, it also carries a considerable financial and social harm on their family members and communities. @traumaLMIC

Ethical considerations
----------------------
The data has been thoroughly anonymized. Each patient have been given an ID number, time frames have been shifted to make identification by admission and discharge or death dates impossible, and a further 47 patients have been omitted from the data set to shift the data frames and disable tracking.

The study is observational, the care provided to each patient has been evidence-based best available care. The data are standard details that would have been recorded anyway, and collection has not delayed or interfered with care nor added any cost.

FRÅGA: jag kan inte se att consent tagits?

Aim
-----------
Predict the unmet need for ICU beds for trauma patients at tertiary hospitals in urban India by applying propensity score matching.

The theory being examined in this thesis is that there exist subset of the TITCO1 dataset that didn't go to the ICU, but are very similar to the patients that did go to the ICU, in the covariates that caused ICU-treated patients to receive this intervention. The thesis is that these patients were not admitted to the IC for lack of such resources. These patients will be identified using propensity score matching.

The size of this group in relation to the group that were admitted to the ICU indicate the magnitude of the unmet need of ICU beds. If, for every patient that did receive ICU care, there was another patient with similar predictors that wasn’t admitted due to scarcity of ICU resources, that would mean the ICU resources are 50% of what they ought to be.

Methods 
=======

Study design
------------

<!-- Present key elements of study design -->
Observational retrospective cohort study from data collected at four tertiary hospitals in urban India.

Setting
-------

<!-- Describe the setting, locations, and relevant dates, including
periods of recruitment, exposure, follow-up, and data collection -->
FRÅGA: kopierat rakt av härifrån https://github.com/titco/titco-I
Men det är ju exakt vad som ska under den här rubriken. Får man göra så eller är det plagiat?
The TITCO-I v1 dataset includes data on 16 000 patients admitted to four public university hospitals in urban India. The original dataset included 16 047 patients but 47 patients were deleted from the dataset for anonymization purposes, see the anonymization section for more details. At the time of the research Jai Prakash Narayan Apex Trauma Center (JPNATC), All India Institute of Medical Sciences, New Delhi, was a dedicated trauma centre with almost 180 beds. King Edward Memorial hospital (KEM), Mumbai, was a tertiary level hospital with no dedicated trauma ward. Lokmanya Tilak Municipal General Hospital (LTMGH), Mumbai was a tertiary level public university hospital with a dedicated trauma ward with 14 beds. The Institute of Post-Graduate Medical Education and Research and Seth Sukhlal Karnani Memorial Hospital (SSKM), Kolkata was a tertiary level public university hospital, but without a dedicated trauma ward. The cost of care in these hospitals was nominal. The first patient was enrolled in July 2013 and the last patient in December 2015. The exact enrollment period varied across participating hospitals.

Participants
------------

<!-- Cohort study: Give the eligibility criteria, and the sources and
methods of selection of participants. Describe methods of
follow-up. For matched studies, give matching criteria and number of
exposed and unexposed -->
The TITCO1 data set includes all data recorded for trauma patients of the emergency departments at the participating hospitals. Trauma covers injury and with a mechanism of road traffic, railway, fall, burn or assault. 
The patients represent mainly the lower stratum of socioeconomic stratum of their communities.

47 patients were excluded for anonymization purposes, 120 patients were excluded because they were dead on arrival, and further XX pat were excluded because of missing data.

<!-- Case-control study: Give the eligibility criteria, and the
sources and + methods of case ascertainment and control
selection. Give the rationale for the choice of cases and
controls. For matched studies, give matching criteria and the number
of controls per case -->

<!-- Cross-sectional study: Give the eligibility criteria, and the
sources and methods of selection of participants -->

Variables and data sources/measurements
---------------------------------------

<!-- Clearly define all outcomes, exposures, predictors, potential
confounders, and effect modifiers. Give diagnostic criteria, if
applicable. For each variable of interest, give sources of data and
details of methods of assessment (measurement). Describe comparability
of assessment methods if there is more than one group -->
There are 193 possible observations recorded for each patient. A general presentation of these observations is in table_nums('tab_1')

```{r}
setwd("C:/Users/maria/Downloads/Packt Learning RStudio for R Statistical Computing 2012 RETAIL eBook-repackb00k/unmet-ICU-beds")
load("table1_1.rdata")

knitr::kable(table1_1, caption = "Data types collected in the TITCO1 dataset")
```

datatvätt:
FRÅGA: quantitativa variablerna har många NA som jag tror egentligen är NO, ex om pat har haft syrgas vid spO2 mätning. kan jag tänka lite fritt här när det känns uppenbart eller ska jag förankra? också finns det ex obs där numeriska värdet för spO2 saknas men ändå data yes/no om pat fått syrgas och jag vet inte om sån data är meningsfull?
FRÅGA: vad betyder GCS verbal V1?  det är inte felregistrerat det är ett alternativ i codebook. jag har hittat ett förslag om att det är pat som inte kan ge verbalt svar pga intub eller surgical airway, men det finns pat som har varken eller och GCS verbalt = 1v

The quantitative data include recorded parameters such as vital parameters, lab results, GCS total score and cumulative time in surgery or ICU, if applicable. Measurement or assessment techniques are standardized.

The qualitative data include categorical data such as coarse categorization or the accident mechanism and injury, details about arrival circumstances, and whether or not a patient had received a medical examination or intervention.

Text data are freehand examination results and operation categorizations, as well as assigned ICD codes. These data are vast and contain a lot of useful information, but also introduce a lot of challenges. There is a lack of standardization, user dependence, heterogenous grouping caused by using too general ICD codes, and datamining challenges for this type of data. If it may cause inaccuracies or practical challenges, it shall be omitted.

Bias 
----

<!-- Describe any efforts to address potential sources of bias -->
The TITCO1 dataset are observational data collected from a clinical environment and not an experimental setting.  Medical interventions, examinations and ICU admission are not randomized, they are a consequence of a patients medical condition. There exists a causal effect between the independent and the dependent variables: a patient is sent to the ICU precisely because of the severity of their condition.

If the purpose of this thesis were to examine the efficacy of ICU care on eg. mortality, this bias would have to be addressed using some statistical tool, but it is not. The purpose of this thesis is use these same methods to assemble a control group but not for the purpose of examining the efficacy of ICU care on eg. mortality, but to identify and quantify the subset of patients that should have been admitted to the ICU but weren't.

Propensity Score Matching will be described in detail in chapter Statistical methods.

FRÅGA: Bör jag tänka på någon bias iallafall? Jag kan tänka mig att det finns skillnader mellan sjukhusen som bör undersökas (de har väl inte nödvändigtvis exakt samma proportioner ICU-sängar givet upptagningsområde, och att ett sjukhus därför kan vara mycket mer generös med ICU-sängarna än ett annat?), men vad för typ av statistisk bias är det isåfall? 
Det finns en selection bias i att det är lägre socioekonomisk stratum, men hur relevant är den?

Propensity score matching in itself introduces a source of bias. The researcher chooses the predictor variables and so can introduce a self-fulfilling prophecy. For this reason the dataset shall be randomly split into two parts of 40% and 60% of the patients. The 40% shall be used to produce the propensity score matching, and the 60% shall be used to verify it. 
Fråga: typ?

Study size 
----------

<!-- Explain how the study size was arrived at -->
The data of 16047 trauma patients were collected at the participating hospitals over 2.5 years, as described in chapter Settings.
47 entries have been randomly excluded for anonymization purposes, and for the purpose of this study the 120 patients who were dead on arrival have been excluded aswell.

Återkom: XX patients were discarded because of missing data

Quantitative variables
----------------------

<!-- Explain how quantitative variables were handled in the
analyses. If applicable, describe which groupings were chosen and why
-->
Fråga: hur diffar detta från vad som finns i Variables and data sources/measurements

Statistical methods
-------------------

<!--

(a) Describe all statistical methods, including those used to control
for confounding
-->
Propensity score matching will be used in this analysis. Propensity score matching (PSM) is best explained using one typical application: identifying a control group when examining the efficacy of an intervention for a specific outcome, where there is causality between the independent variables for the intervention, and sometimes also the outcome. There is a need to offset or balance the effect of this causality, but a randomized case-control study could be impractical or immoral.

In this context, it could be to examine the efficacy of ICU care on survival. If the correlation between the independent variables and the intervention and outcome isn't accounted for, a simple comparison between the cohort that did go to the ICU to the cohort that did not, would find that ICU care actually increased mortality. In the TITCO1 dataset the survival rate of patients who did not go to the ICU was 81% versus 68% for patients who did. This is of course an incorrect conclusion, this data has treatment selection bias as patient covariates often confound treatment selection. The higher mortality observed in patients who received ICU care is because patients were selected for the intervention because they presented certain medical conditions that also carry a high mortality - even with the intervention ICU care. Producing a randomized control trial study for these types of lifesaving medical interventions for trauma patients would obviously go against ethical guidelines of research, and another approach to produce a control group is necessary. Propensity score matching is a statistical procedure for reducing this bias by assembling a sample from within the data set, in which confounding factors are balanced between the cohort that got the intervention and the cohort that did not. 
<!-- 
The goal of propensity score matching is usually to identify suitable control subjects that did not receive an intervention, even though they fairly well met the same inclusion criteria that the subjects that did receive the intervention had. Control subjects who are similar enough to the intervention group in the most relevant aspects for receiving treatment. 
-->
The size of this subset in relation to the subset of patients that did receive ICU care is the end point for this analysis. This proportion is indicative of the unmet need of ICU beds, as described in chapter Aim.
In the typical PSM appliction, this subset is used as the control group to determine the efficacy of ICU care for some outcome, eg mortality.

The first step is to identify predictor variables: covariates most predictive of the dependent variable. In this context it means to first identify the conditions and medical interventions received that correlate best to being offered ICU care. All properties of the patients that were sent to the ICU will be compared to the patents who didn't, in order to identify properties that distinguishes this cohort.

Second step is to calculate propensity scores for all patients. Each patient in the TITCO1 data set gets a propensity score based on their medical conditions and interventions, specifically their individual values of the predictor variables. The propensity score determines the likelihood that a patient will receive ICU care, given their condition.

Third step is PSM. PSM is a standardized mathematical approach to identify members of the subset that did not get the intervention but had the same predictors as the subgroup that did get the intervention. The purpose of the propensity score matching is identifying patients that weren’t admitted into the ICU, but presented with a set of predictors - a propensity score - very similar to those that were admitted.

(b) Describe any methods used to examine subgroups and interactions

The two subgroups were patients admitted to the ICU, and the patients who weren't. Admission to the iCU is a binary yes or no, length of stay has not been taken into Consideration.

Missing Data
-------------
The data completion rate is presented in table_nums('tab_1'). As the base assumption, all quantitative variables collected for each patient are considered potential recordings. Data that are missing are assumed missing because care for a patient did not motivate the collection of the data. For example, if serum creatinine was not recorded, it is assumed that the collection of this variable wasn't relevant for care of the patient.

There are some variables that should reasonably have been collected and documented given the patient condition, but omission will not be cause of dismissal unless these variables are predictors.

All patients with incomplete observations for any predictors will be discarded. For relevant qualitative variables, missing data did not automatically warrant dismissal, if the missing data could be indirectly proven. For example if the quantitative yes/no for an examination like a CT scan was missing, it was deemed a 'yes' if there was a CT scan result recorded. If both were missing, the patient data was dismissed.
All vital parameters are considered mandatory, irregardless of patient condition so patients without data of heart rate, systolic blood pressure or respiration rate at any time of their hospital stay, have been discarded.  
FRÅGA: right?

All quantitative data from the original dataset that were entered as zero but where a zero value is physiologically impossible were replaced with NA values. Zero values for variables where it is possible to have zero values, such as respiration rate and heart rate were left as zeroes to not tamper with the data, although a zero value indicate that the variable probably wasn't measured. These amount to 250 and 155 entries respectively for the first observation within an hour and less than 100 for both for the second observation after 24 h, so the choice to include or exclude these observations won’t skew the data considerably.

(d)

Cohort study: If applicable, explain how loss to follow-up was addressed

Case-control study:If applicable, explain how matching of cases and
controls was addressed

Cross-sectional study: If applicable, describe analytical methods
taking account of sampling strategy

(e) Describe any sensitivity analyses
-->

Predictor selection 
-------------------
The quantitative data did not produce any clear results see table_nums('tab_2')

```{r}
setwd("C:/Users/maria/Downloads/Packt Learning RStudio for R Statistical Computing 2012 RETAIL eBook-repackb00k/unmet-ICU-beds")
load("table3.rdata")

knitr::kable(table3, caption = "Means and standard devations of quantitative variables of ICU and non-ICU admitted patients, respectively")
```

This is perhaps because of a crowd out effect. For instance, the total Glascow coma scale score is high and quite similar for the non-ICU subset versus the ICU subset, the difference between the average of the groups is 2 points. However, considered individually, low scores in either category is a high propensity to admit a patient to the ICU, as demonstrated in the comparison of quantitative parameters.

P-values and T-values of non-ICU vs ICU patients:

```{r}
setwd("C:/Users/maria/Downloads/Packt Learning RStudio for R Statistical Computing 2012 RETAIL eBook-repackb00k/unmet-ICU-beds")
load("statresult.rdata")

knitr::kable(table3, caption = "P- and t-values of the quantitative values in comparison of non-ICU and ICU patients")
```

FRÅGA: är p-värdena rimliga tycker du? de är superlåga och min känsla - med 5 hp statistik i ryggen - är att std-avvikelsen är för stor för det?

For the qualitative data, the propensity of each parameter to get a patient admitted to the ICU is in relation to the proportion of patients that were admitted to the ICU from the whole TITCO1 data set, about a third. If considerably more than a third of each qualitative parameter are admitted to the ICU, then that parameter is of interest. The proportions are presented in

```{r}
setwd("C:/Users/maria/Downloads/Packt Learning RStudio for R Statistical Computing 2012 RETAIL eBook-repackb00k/unmet-ICU-beds")
load("qualitativetable.rdata")

knitr::kable(qualitativetable, caption = "Proportion of patients admitted to the ICU for all variants of all quuantitative variables")
```

Results
=======

<!--

Participants
------------

(a) Report numbers of individuals at each stage of study—eg numbers
potentially eligible, examined for eligibility, confirmed eligible,
included in the study, completing follow-up, and analysed

(b) Give reasons for non-participation at each stage

(c) Consider use of a flow diagram

Descriptive data
----------------

(a) Give characteristics of study participants (eg demographic,
clinical, social) and information on exposures and potential
confounders

(b) Indicate number of participants with missing data for each
variable of interest

(c) Cohort study — Summarise follow-up time (eg, average and total
amount)

Outcome data
------------

Cohort study — Report numbers of outcome events or summary measures
over time

Case-control study — Report numbers in each exposure category, or
summary measures of exposure

Cross-sectional study — Report numbers of outcome events or summary
measures

Main results
------------

(a) Give unadjusted estimates and, if applicable, confounder-adjusted
estimates and their precision (eg, 95% confidence interval). Make
clear which confounders were adjusted for and why they were included

(b) Report category boundaries when continuous variables were
categorized

(c) If relevant, consider translating estimates of relative risk into
absolute risk for a meaningful time period

Other analyses 
--------------

Report other analyses done—eg analyses of subgroups and interactions,
and sensitivity analyses 
-->

You can include code in this document like this:

```{r main, echo=FALSE}
source("main.R") ## This "imports" the main script file of your project and run any code in it
```

You can also embed plots:

```{r plot, echo=FALSE}
plot(pressure)
```

You can also mix text and code, so called inline code, like this: `r 2+5`.

Discussion
==========

<!--

Key results
-----------
Summarise key results with reference to study objectives

Limitations
-----------

Discuss limitations of the study, taking into account sources of
potential bias or imprecision.  Discuss both direction and magnitude
of any potential bias

Interpretation
--------------

Give a cautious overall interpretation of results considering
objectives, limitations, multiplicity of analyses, results from
similar studies, and other relevant evidence

Generalisability
----------------

Discuss the generalisability (external validity) of the study results

-->

Conclusion
==========

<!-- Give a short conclusion, corresponding to the aim -->

References
==========

<!-- Do not edit by hand, references will be inserted and formatted automatically once you knit this document -->