### 2022-09-22 manuscript.Rmd
intermediär version

---
title: 'Quantifying unmet ICU beds in tertiary hospitals in urban India using Propensity
  Score Matching '
author: "Maria Eriksson"
output:
  html_document: default
  word_document: default
bibliography: bibliography.bib
csl: vancouver.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
table_nums <- captioner::captioner(prefix = "Tab.")
```

<!-- This is an R Markdown document. Markdown is a simple formatting
syntax for authoring HTML, PDF, and MS Word documents. For more
details on using R Markdown see http://rmarkdown.rstudio.com -->

<!-- This is a comment and will not be present in the compiled
document. Try "knitting" this document by clicking the knit button up
to the left, or by running `rmarkdown::render("manuscript.Rmd")` in
the console and you should only see headings -->

<!-- Structure and most comments are from the Strengthening the
reporting of observational studies in epidemiology (STROBE) statement,
see https://www.strobe-statement.org/. For more explanations see the
paper Strengthening the Reporting of Observational Studies in
Epidemiology (STROBE):explanation and elaboration.  -->

Abbreviations
============
ACS - American college of surgeons
ER - Emergency room
HIC - High-income countries
LMIC - Low- and middle-income countries
PSM - propensity score matching
RCT - Randomized case-control study
TITCO -Towards Improved Trauma Care Outcomes 
TQIP - Trauma Care Improvement Program

<!-- Indicate the study’s design with a commonly used term in the
title, for example: "Associations between resuscitation interventions
and opportunities for improvement in adult trauma patients: A cohort
study" -->

Abstract
========

<!-- Provide in the abstract an informative and balanced summary of
what was done and what was found. Not more than 300 words. -->

Background
----------

Methods
-------

Results
-------

Conclusion
----------

Introduction
============

<!-- Explain the scientific background and rationale for the
investigation being reported. End by stating the aim or specific
objectives, including any prespecified hypotheses. The introduction
should not be longer than 1-1.5 pages, typically 4-5 paragraphs, each
6-8 lines long. Support every statement with a citation -->

Trauma globally and in India
----------------------------
Trauma is a substantial global health problem that disproportionately affects low and middle-income countries (LMIC), resulting in significant mortality and morbidity @traumaLMIC. Despite having only 1% of the global vehicle fleet, a disproportionate 10% of global traffic casualties occur in India @indiantrauma. The annual mortality from injury in India amount to one million casualties @onemilliondead. Half of trauma mortality in India occur in-hospital, and in-hospital trauma mortality in Indian hospitals was double that of high-income countries @indiantrauma.

Mortality due to accidents in India amounted to 421 104 people in 2019, which is 31.5 per 100.000 inhabitants. The traumatic mechanisms of injury from accidents covered in this thesis are road accidents, falls, assault and burns and these make up about half of the accident mortality @nationalcrime.

Trauma is a main cause of mortality and morbidity to otherwise healthy and young individuals, and dispropotionately affect this demographic. Most casualties affected the age group 30-45 years, followed by the age bracket 18-30 years @nationalcrime. People in this demographic are also oftentimes the primary or sole breadwinners for their families. Therefore the detrimental impact of trauma go beyond mortality or substantial loss of disease-free years of life for the affected individuals, it also carries a considerable financial and social harm on their family members and communities @traumaLMIC @nationalcrime

Health care professionals in many LMIC testify that their medical assessment oftentimes is that a trauma patient should be admitted to the ICU, but are not because such resources are scarce, overloaded and unavailable. Their medical judgement is that the lives of these patients were salvageable, and that these patients died as a direct result of not being admitted to the ICU @stuffLMICdoctorssay. Globally, an estimated two million lives could be saved annually, if the appropriate and timely trauma care provided by hospitals in many high-income countries were available and provided in low to-middle-income countries (LMICs) @livessaved

A diverse mix of public, for-profit and not-for-profit private actors provide health services in India. According to a WHO 2022 report, India has 76 709 non-ventilator supported and 39 476 ventilator-supported ICU beds @ICUbeds. This is 8.2 per 100 000 inhabitants.
The WHO report also highlight the probable shortage of ICU resources in India, though in the context of the COVID-19 pandemic.

Methods to examine trauma care and ensure equivalence
-----------------------------------------------------
In order to improve and ensure equivalence of intra-hospital care, trauma quality improvement programs (TQIP) have been developed. The pilot and most notable example is the TQIP developed by the American College of Surgeons ACS. Since the implementation of the ACS pilot, TQIP has evolved to a broad term that include many kinds of processes to improve the quality of trauma care, such as mortality and morbidity conferences, using audit filters or comparing risk adjusted outcomes over time and between centres. The aim is to find discrepancies in treatment for patients presenting with similar conditions. TQIPs aim to ensure patients get equivalent care within and across hospitals, to benchmark performance of health care providers versus their peers and to improve trauma care by developing best practice guidelines @tqip.

One aspect of TQIP assessments is to collect trauma data from participating health care providers and apply statistical analysis to the data. From this data, TQIPs utilizes a retrospective cohort study, analysing the current of trauma care patients and post-fact attempting to determine the appropriate level of care for each patient. From these analyses many aspects of trauma care can be improved and optimized, such as determining the appropriate amount of ICU beds and other trauma care resources @tqip.

It is tempting to apply existing models of TQIP assessments on Indian trauma data to identify causes of the observed disparities in trauma care performance in HIC and LMIC, and propose measures to harmonize performance. Unfortunately as TQIP often originate in high-income countries (HIC), they are either not applicable nor easily transferred to the hospital conditions of LMIC @frailty @gunshot  @onemilliondead. Applying TQIPs outside of their original context or even within a subset of its own origin is difficult even without taking income bracket in account, for instance firearm violence is a considerable bigger portion of trauma patients in some US contexts, than the national average @gunshot. For trauma patients, there exist significant structural differences between HIC and LMIC. These differences are in factors such as injury mechanisms and severity, health care resources as well as patient make-up @frailty @indiantrauma @traumaLMIC @onemilliondead. One example of this is the differences in helmet use for motorcyclists and their passengers, as well as differences of displacement in vehicle fleets in LMIC versus HIC. These factors have tremendous impact on the type and severity of injuries patients in the ER have, and subsequently on their care and outcome. These differences make a direct transferral of HIC TQIP models onto LMIC irrelevant or misleading and there is a need to develop TQIP assessment applicable to local conditions.

Kommentar på Mgerdin kommentar: jag snodde din definition här. Jag håller med om att detta inte är den mest relevanta informationen men för "kontrollmekanism"-arbetet som ska in på måndag här har KI en checklista och de vill ha 3-5 sidor bakgrund trots att vi jobbar med max 1-1.5 så jag måste feta ut. Detsamma för kommentaren i Aim jag är medveten om att det är felplacerat eller irrelevant.

Knowlege gap
------------
There are compelling reasons to believe there is a shortage of ICU beds in India, but the scope of this shortage and the potential for reduction of trauma mortality if this shortage would be remedied is unknown.

WHO and local health care workers believe there is a shortage of ICU beds in Indian hospitals, and that the lives of many patients would have been salvageable if the appropriate ICU resources had been available. The differences of in-hospital trauma mortality in Indian hospitals was double that of high-income countriesin, and part of this difference could be explained by a shortage of ICU beds in India @indiantrauma

Aim
----

Quantify the unmet need for ICU beds in tertiaty hospitals in urban India.

Apply statistical analysis to a large data set of Indian trauma patients and determine how many out of the patients that were not admitted to the ICU reasonably should have been admitted given their condition. Their numbers in proportion to the number of ICU-admitted patients is indicative of the relative shortage of ICU beds.

Methods 
=======

Study design
------------

<!-- Present key elements of study design -->
Observational retrospective cohort study using data from the Towards Improved Trauma Care Outcomes (TITCO) in India cohort.

Setting
-------

<!-- Describe the setting, locations, and relevant dates, including
periods of recruitment, exposure, follow-up, and data collection -->
16 000 patients admitted to four public university hospitals in urban India are included in the TITCO dataset @titco. Initially the data set included 16047 patients, but to improve anonymization 47 patients were deleted from the dataset as described in the chapter anonymization. The participating hospitals were King Edward Memorial hospital (KEM) and Lokmanya Tilak Municipal General Hospital (LTMGH), Mumbai, Jai Prakash Narayan Apex Trauma Center (JPNATC), All India Institute of Medical Sciences, New Delhi and The Institute of Post-Graduate Medical Education and Research and Seth Sukhlal Karnani Memorial Hospital (SSKM), Kolkata. When the data was collected all hospitals were tertiary hospitals. JPNATC had almost 180 beds in a dedicated trauma centre, KEM had no specific trauma ward, LTMGH was a public university hospital with a specific trauma ward with 14 beds and SSKM was a  public university hospital with no specific trauma ward.

The participating hospitals are all public entities, so a majority of patients came from low socioeconomic circumstances in their respective communities.

The data was collected between July 2013 and December 2015 but the collection period for any individual hospitals did not cover necessarily this whole period.

A dedicated project officer involved in the data collection was on site five days per week, approximately eight hours a day. The project officers covered night, evening and morning shifts in a way so all shift variants were covered every month of collection. Project officers worked in the area where trauma patients arrived to the hospital and collected data for patients arriving during their shift through direct observation. If the officer observed an examination being performed but not documented in the patient records were allowed to ask health care staff for the result. If a trauma patient arrived when a project officer was not on site, they were identified from the log books of the nurses on staff in the emergency department and their data was collected from patient records.

Participants
------------

<!-- Cohort study: Give the eligibility criteria, and the sources and
methods of selection of participants. Describe methods of
follow-up. For matched studies, give matching criteria and number of
exposed and unexposed -->
The eligibility criteria for TITCO were patients who presented with injuries from trauma at the emergency departments of the participating hospitals with a mechanism of road traffic, railway, fall, burn or assault. Patients who were dead on arrival were excluded, as were patients who only had isolated limb injuries.

<!-- Case-control study: Give the eligibility criteria, and the
sources and + methods of case ascertainment and control
selection. Give the rationale for the choice of cases and
controls. For matched studies, give matching criteria and the number
of controls per case -->

<!-- Cross-sectional study: Give the eligibility criteria, and the
sources and methods of selection of participants -->
Ethical considerations
----------------------

This study does not requlre ethical permission, it refers to data collected and shared openly.

Ethical approval for the collection of the data used was obtained from the boards of ethical approval in all four hospitals.

SSKM, LTMGH, KEM and JPNATC name of ethical review boards and approval reference numbers were IPGME&R Research Oversight Committee (IEC/279), Ethics Committee of the Staff and Research Society (IEC/11/13), Institutional Ethics Committee (IEC(I)/OUT/222/14), Institute Ethics Committee All India Institute of Medical Sciences (EC/NP-279/2013 RP-Ol/2013), respectively.

The ethical review boards of all hospitals granted a waiver of informed consent. The trauma patients of in this study included patients in poor physiological, psychological and physical state, oftentimes unconscious or with reduced level of consciousness. The data collected for this study were merely routine data that would have been collected under normal circumstances and did not involve any additional examinations. The collection did not in any way interfere, delay or alter the care provided. Therefore the research team and the ethical review boards concluded that obtaining informed consent would only unneccesarily burden patients and/or family members. Personal details such as names, insurance details or contact information were not recorded.

Anonymization
-------------
Hospital identification number
Hospital name has been replaced with a identification number, and the conversion key has not been saved so there is not possible to reverse the conversion. Despite this, a high degree of anonymization can not be expected because the hospitals do have some distinguishable features and these are reflected in the data collected. For example,  70 % of the railroad accidents in the data set is in one of the hospitals, and one of the hosapitals is in close proximity to a big railroad hub.

Patient identification numbers
Patient names were never recorded, patients were identified by a identification number in the original unencrypted data set. In the anonymization process the identification number has been replaced with another random identification number, and the conversion key has not been saved so there is not possible to reverse the conversion.

Date observations
Dates recorded in the TITCO-I data set were shifted using a random offset of somewhere between 0 and 5 weeks for each patient, but without losing any relevant information. The offset was designed to not dissipate details that could be of potential use for researchers such as day of week and rough indication of season. The internal relations between dates have been preserved so that the timing of all recordings and discharge in chronological relation to admission are all intact.

It is important to note that all patients have sequence numbers that corresponded to the chronological order in which they arrived at any of the participating hospitals, and that sequence number has been kept intact. Keeping this sequence number variable intact allows for many useful operations such as examinations of trends in the data, while it is not possible to undo the date randomization by the sequence number variable, such an equation system is still underdetermined. Unfortunately this means it is also not possible to reconstruct the patient volumes at any given hospital on any given day and from that rate of resource engagement. Nor is it possible to distinguish if patients who were made to wait for their first assessment were so because their injuries were deemed as less severe or if it was because health care staff was overwhelmed which is an important distinction.

In addition, 47 patient data were randomly deleted to prevent identification via the sequence number variable.

Age
Very old patients are few, they have been grouped together to prevent identification via the age parameter.

Text based input
1% of the text-based inputs in ICD code fields and examination or operation findings were randomly deleted to obstruct identification by this variable.

Variables and data sources/measurements
---------------------------------------

<!-- Clearly define all outcomes, exposures, predictors, potential
confounders, and effect modifiers. Give diagnostic criteria, if
applicable. For each variable of interest, give sources of data and
details of methods of assessment (measurement). Describe comparability
of assessment methods if there is more than one group -->
There are 193 possible observations recorded for each patient. A general presentation of these observations is in table_nums('tab_1')

```{r}
setwd("C:/Users/maria/Downloads/Packt Learning RStudio for R Statistical Computing 2012 RETAIL eBook-repackb00k/unmet-ICU-beds")
load("table1_1.rdata")

knitr::kable(table1_1, caption = "Data types collected in the TITCO-I dataset")
```

datatvätt:
FRÅGA: quantitativa variablerna har många NA som jag tror egentligen är NO, ex om pat har haft syrgas vid spO2 mätning. kan jag tänka lite fritt här när det känns uppenbart eller ska jag förankra? också finns det ex obs där numeriska värdet för spO2 saknas men ändå data yes/no om pat fått syrgas och jag vet inte om sån data är meningsfull?

The quantitative data include recorded parameters such as vital parameters, lab results, GCS total score and cumulative time in surgery or ICU, if applicable. Measurement or assessment techniques are standardized.

The qualitative data include categorical data such as coarse categorization or the accident mechanism and injury, details about arrival circumstances, and whether or not a patient had received a medical examination or intervention.

Text data are freehand examination results and operation categorizations, as well as assigned ICD codes. These data are vast and contain a lot of useful information, but also introduce a lot of challenges. There is a lack of standardization, user dependence, heterogenous grouping caused by using too general ICD codes, and datamining challenges for this type of data. If it may cause inaccuracies or practical challenges, it shall be omitted.

Bias 
----

<!-- Describe any efforts to address potential sources of bias -->
The TITCO dataset are observational data collected from a clinical environment and not an experimental setting.  Medical interventions, examinations and ICU admission are not randomized, they are a consequence of a patients medical condition. There exists a causal effect between the independent and the dependent variables: a patient is sent to the ICU precisely because of the severity of their condition.

There exist a potential bias in the routines, resources and patient makeup of the participating hospitals. Becasue they are all tertiary hospitals, about two thirds of the patients have been referred from other health care providers, meaning that there is a selection bias in the patient populations. Further, hospitals with dedicated trauma wards have a propensity to receive transferred patients or police escorted patients with worse condition and prognosis.

As stated in chapter Settings, most patients have low socioeconomic status. Low socioeconomic status is associated with higher level of malnutrition and comorbidities and would have a negtive effect on vitals and the capacity of patients to tolerate certain interventions. This circumstance doesn't affect the analysis and internal validity, but reduce the generalizability of the results. 
<!-- (ÅTERKOM: har inte råd att refereas?) -->

Because the purpose of this thesis is to statistically assess how many patients that should have gone to the ICU but did not, these particular selection biases are not expected to skew the outcome, and therefore does not have to be adressed.

Hospitals with trauma wards could have lower admission criteria to the ICU than hospitals with no trauma ward. Because examining admission criteria to the ICU is part of the analysis it may be necessary to examine and address this bias by comparing ICU admitted patients between hospitals. Propensity score matching can be used to identify differences in ICU admission criteria between hospitals and find patients that were admitted but need not have been. These patients are excluded from predictor selection to reduce this bias. This is part of the development and training of the PSM model.

Propensity score matching in itself introduces a source of bias from unfit assumptions. The terminology and potential sources of bias are explained chapter Study Design. In choosing predictor variables the researcher could introduce an overdetermined set and a self-fulfilling prophecy. For this reason the dataset shall be randomly split into two parts of 40% and 60% of the patients. The 60% shall be used to produce the propensity score matching, and the remaining 40% shall be used to verify it.

Study size 
----------

<!-- Explain how the study size was arrived at -->
The data of 16000 trauma patients were collected at the participating hospitals over 2.5 years, as described in chapter Settings.

Statistical methods
-------------------
<!--

(a) Describe all statistical methods, including those used to control
for confounding
-->
Propensity score matching (PSM) is useful method for identifying and assembling a control group from within a data set where there is causality between the independent and dependent variables, and a randomized case-control (RCT) study to offset or balance this bias is impractical, immoral or otherwise improper. Propensity score is a statistical variable, understood as the likelihood of a subject receiving treatment conditional on observations of relevant covariates of that subject before treatment. In RCTs it is known from study design, but it is generally unknown in observational data. The rationale for PSM is that while it is unknown, this variable can be fairly well estimated using logistic regression.

The theory being examined in this masters thesis is that there exist a subset of the TITCO dataset that did not go to the ICU, but are very similar to the patients that did go to the ICU, in the covariates that caused ICU-treated patients to receive this intervention. The thesis is that these patients were not admitted to the ICU only for lack of such resources, not by a medical assessment.

Translated into this context this means that in this data set exist a selection bias, there is causality between the independent variables medical conditions and interventions and the dependent variable ICU admittance. Patients were admitted to the ICU as a consequence of their medical condition. Performing a randomized case-control study and randomly assigning or withholding a live-saving treatment would obviously be immoral, but we believe that in the TITCO data set exist patients who were not admitted into the ICU by volition but for scarcity of such resources. Such a group would be akin to an non-experimental control group and PSM can be used to identify and assemble it.  If such a subset can be identified by PSM, then it naturally follows that there exists shortage of ICU beds in tertiary hospitals in India. It also follows that the shortage can be quantified by the proportionality of this subset versus the subset that did receive ICU care. If, for every patient that did receive ICU care, there was another patient with similar predictors that wasn’t admitted due to scarcity of ICU resources, that would mean the ICU resources are 50% of what they ought to be.

The propensity score model is developed using the Training Data and tested using the Validation Data. The workflow contain a series of steps and the work flow should understood as linear, but iterative. The output of each step is the input for later steps. Therefore the output of the last step reflect the cumulative error from unfit assumptions made the second and fourth step. The output is tested against Validation Data. By reverting back and adjusting assumptions in an iterative process, the model is continuously improved and the accuracy and validity is improved.

The first step of PSM is to randomly split the data set into two parts, 60% and 40%. The 60% part is a Training Part used to delvelop and train the PSM, and the 40% part is the Validation Part. This is done because the inherent bias of the PSM method described in chapter Bias. Training Data is used in steps 2-4 and the output is then validated using Validation Data.

Second step is to identify predictor variables: covariates most predictive of the dependent variable, as described in chapter Predictor selection. In this context it means to first identify the conditions and medical interventions received that correlate best to being admitted to ICU care. This selection is based on clinical experience as well as analysis of the Training Data. In the data analysis all properties of the patients that were sent to the ICU will be compared to the patents who did not, in order to identify properties that distinguishes this cohort., Familiarity with the collection of source data is crucial to minimize the risk to choose irrelevant predictors and introduce error.

Third step is to build a standard multivariable logistic regression model with ICU care as the outcome and the predictors selected in the second step as independent variables. This model is used to calculate the standard variable propensity score for all patients. The propensity score determines the likelihood that a patient will receive ICU care, given their condition.

Fourth step is PSM. In PSM members of the subset that did not get ICU care but had similar predictors - and therefore similar propensity scores - as the subgroup that did get ICU care, are matched to patients who did go to the ICU. There are a few different ways to perform this matching but they are all imperfect because propensity scores are unique and perfect matching is unattainable There are control mechanisms and limitations but a risk of applying too fix or too loose matching criteria the outcome could inadvertedly vastly under- or overestimate the  need of ICU beds remain.

The same procedure is repeated on the Validation Data using the same predictors and the same matching criteria, and this should produce a cohort comparable to the one trom Training Data.

(b) Describe any methods used to examine subgroups and interactions

The two subgroups were patients admitted to the ICU, and the patients who were not. Admission to the iCU is a binary yes or no, length of stay has not been taken into Consideration.

Missing Data
-------------
The data completion rate is presented in table_nums('tab_1'). As the base assumption, all quantitative variables collected for each patient are considered potential recordings. Data that are missing are assumed missing because care for a patient did not motivate the collection of the data. For example, if serum creatinine was not recorded, it is assumed that the collection of this variable was not relevant for care of the patient.

There are some variables that should reasonably have been collected and documented given the patient condition, but omission will not be cause of dismissal unless these variables are predictors.

All patients with incomplete observations for any predictors will be discarded. For relevant qualitative variables, missing data did not automatically warrant dismissal, if the missing data could be indirectly proven. For example if the quantitative yes/no for an examination like a CT scan was missing, it was deemed a 'yes' if there was a CT scan result recorded. If both were missing, the patient data was dismissed.
All vital parameters are considered mandatory, irregardless of patient condition so patients without data of heart rate, systolic blood pressure or respiration rate at any time of their hospital stay, have been discarded.  
FRÅGA: right?

All quantitative data from the original dataset that were entered as zero but where a zero value is physiologically impossible were replaced with NA values. Zero values for variables where it is possible to have zero values, such as respiration rate and heart rate were left as zeroes to not tamper with the data, although a zero value indicate that the variable probably was not measured. These amount to 250 and 155 entries respectively for the first observation within an hour and less than 100 for both for the second observation after 24 h, so the choice to include or exclude these observations won’t skew the data considerably.

(d)

Cohort study: If applicable, explain how loss to follow-up was addressed

Case-control study:If applicable, explain how matching of cases and
controls was addressed

Cross-sectional study: If applicable, describe analytical methods
taking account of sampling strategy

(e) Describe any sensitivity analyses
-->

Predictor selection 
-------------------

Predictor variables will be selected using both clinical judgement and data analysis.

Clinical experience compels inclusion of vital parameters, irregardless if there is any statistical motivation. Such variables are heart rate, systolic blood pressure and respiration rate.

The TITCO-I data set is analyzed mathemathically in pursuit of suitable predictor variables. The data set is divided into two subgroups, the cohort that were admitted to the ICU and the cohort that was not. In comparison between these subgroups, identify variables that distinguishes the former from the latter. The quantitative and qualitative properties were considered separately.

Mean and standard deviation for the quantitative data is analysed for the ICU and non-ICU cohort respectively. p- and t-values when comparing these cohorts to show if there are any statistically significant differences.

For the qualitative data, the propensity of each parameter to get a patient admitted to the ICU is in relation to the proportion of patients that were admitted to the ICU from the whole TITCO-I data set, about a third. If considerably more than a third of each qualitative parameter are admitted to the ICU, then that parameter is of interest.

Some qualitative parameters such as intubation are recorded after both 1 and 24 hours, if applicable. The record at 1 hour has two options indicating the patient is intubated, "before arrival" meaning the patient was intubated at a their first health care facility and "yes" meaning the patient was intubated within the first hour. Because these two options both mean the patient was intubated after 1 hour, an aggregate parameter has been created for this, and a further parameter "intubated after 24 hours" that summarize the patients who were intubated on arrival, the patients who were intubated within the first hour and the patients who were intubated between 1 and 24 hours.

FRÅGA: det finns pat som är intuberade "before arrival" men som inte är noterade som transferred från ett annat sjukhus. De har kommit med motor rickshaw. Är det rimligt att anta att de är transferred från ett sjukhus iallafall för vem har annars intuberat dessa patienter?

<!-- 
The parameter "taken to the operation theatre" is a heterogenous qualitative parameter, where only a small proportion of patients had been taken within the first hour but a substantial proportion of patients had been taken to the operation theatre after 24 hours. There is a need to consider the free text field operation type to filter out for operation types with higher propensity for ICU admission from operations with lower propensity. 
-->
The parameter "taken to the operation theatre" is a heterogenous qualitative parameter, and operations performed in the first hour tend to be more acute than operations performed within the first day of hospitalization. There is a need to consider the free text field operation type to filter out for operation types with higher propensity for ICU admission from operations with lower propensity. 
FRÅGA: borde man ta operation within 24 hours och komplettera med text based operationstyp?

The free hand text based information available contain information about examination findings, surgery types and ICD codes. Many of these data fields contain crucial information about patient condition, but also present a few challenges as described in chapter: Variables and data sources/measurements. These text based data fields will be examined for entries that are more prevalent in the ICU cohort than the non-ICU cohort, and combined with clinical experience to select suitable predictors.

Results
=======

Missing data
------------
Återkom: XX patients were discarded because of missing data

Predictor selection
-------------------
The quantitative data results see table_nums('tab_2')

```{r}
setwd("C:/Users/maria/Downloads/Packt Learning RStudio for R Statistical Computing 2012 RETAIL eBook-repackb00k/unmet-ICU-beds")
load("table3.rdata")

knitr::kable(table3, caption = "Means and standard devations of quantitative variables of ICU and non-ICU admitted patients, respectively")
```

For many parameters the difference is small, perhaps because of a crowd out effect. For instance, the total Glascow coma scale score is high and quite similar for the non-ICU subset versus the ICU subset, the difference between the average of the groups is 2 points. However, low scores in either category is a high propensity to admit a patient to the ICU, as demonstrated in the comparison of quantitative parameters.

P-values and T-values of non-ICU vs ICU patients:
```{r}
setwd("C:/Users/maria/Downloads/Packt Learning RStudio for R Statistical Computing 2012 RETAIL eBook-repackb00k/unmet-ICU-beds")
load("statresult.rdata")

knitr::kable(statresult, caption = "P- and t-values of the quantitative values in comparison of non-ICU and ICU patients")
```
FRÅGA: är p-värdena rimliga tycker du? de är superlåga och min känsla - med 5 hp statistik i ryggen - är att std-avvikelsen är för stor för det?

The proportions of the qualitative data are presented in

```{r}
setwd("C:/Users/maria/Downloads/Packt Learning RStudio for R Statistical Computing 2012 RETAIL eBook-repackb00k/unmet-ICU-beds")
load("qualitativetable.rdata")

knitr::kable(qualitativetable, caption = "Proportion of patients admitted to the ICU for all variants of all quuantitative variables")
```

<!--

Participants
------------

(a) Report numbers of individuals at each stage of study—eg numbers
potentially eligible, examined for eligibility, confirmed eligible,
included in the study, completing follow-up, and analysed

(b) Give reasons for non-participation at each stage

(c) Consider use of a flow diagram

Descriptive data
----------------

(a) Give characteristics of study participants (eg demographic,
clinical, social) and information on exposures and potential
confounders

(b) Indicate number of participants with missing data for each
variable of interest

(c) Cohort study — Summarise follow-up time (eg, average and total
amount)

Outcome data
------------

Cohort study — Report numbers of outcome events or summary measures
over time

Case-control study — Report numbers in each exposure category, or
summary measures of exposure

Cross-sectional study — Report numbers of outcome events or summary
measures

Main results
------------

(a) Give unadjusted estimates and, if applicable, confounder-adjusted
estimates and their precision (eg, 95% confidence interval). Make
clear which confounders were adjusted for and why they were included

(b) Report category boundaries when continuous variables were
categorized

(c) If relevant, consider translating estimates of relative risk into
absolute risk for a meaningful time period

Other analyses 
--------------

Report other analyses done—eg analyses of subgroups and interactions,
and sensitivity analyses 
-->

You can include code in this document like this:

```{r main, echo=FALSE}
source("main.R") ## This "imports" the main script file of your project and run any code in it
```

You can also embed plots:

```{r plot, echo=FALSE}
plot(pressure)
```

You can also mix text and code, so called inline code, like this: `r 2+5`.

Discussion
==========

<!--

Key results
-----------
Summarise key results with reference to study objectives

Limitations
-----------

Discuss limitations of the study, taking into account sources of
potential bias or imprecision.  Discuss both direction and magnitude
of any potential bias
-->

In order to limit the scope of this thesis work, a few quite relevant factors could not be considered. One of the most important factors was temporality. Emergency rooms everywhere face significant variations in patient loads over the week or special occasions such as the Diwali festival. A different approach completely but to the same effect would have been to consider the ICU wards at the participating hospitals and chronologically determine which patients did not receive ICU care because these wards were at capacity. Though intuitively it would seem possible to do given the TITCO-I dataset, it would have been very difficult to do practically.

Length of stay in hours is ICU care was recorded in the TITCO-I data set, but for the purpose of this thesis ICU admission has been reducedn to a binary 'yes' or 'no'. Because it is not uncommon to discontinue ICU care prematurely in favor of patients with more dire needs, the next step would be to consider such circumstances.

<!--
Interpretation
--------------

Give a cautious overall interpretation of results considering
objectives, limitations, multiplicity of analyses, results from
similar studies, and other relevant evidence

Generalisability
----------------

Discuss the generalisability (external validity) of the study results

-->

Conclusion
==========

<!-- Give a short conclusion, corresponding to the aim -->

References
==========

<!-- Do not edit by hand, references will be inserted and formatted automatically once you knit this document -->